---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    out.width = "100%"
)
```

# BiocManager

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/BiocManager)](https://cran.r-project.org/package=BiocManager)
[![Travis build status](https://travis-ci.org/Bioconductor/BiocManager.svg?branch=master)](https://travis-ci.org/Bioconductor/BiocManager)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/Bioconductor/BiocManager?branch=master&svg=true)](https://ci.appveyor.com/project/Bioconductor/BiocManager)
[![CRAN release](http://www.r-pkg.org/badges/version-last-release/BiocManager)](https://github.com/Bioconductor/BiocManager/releases)
[![CRAN downloads](http://cranlogs.r-pkg.org/badges/BiocManager)](https://cran.r-project.org/package=BiocManager)
<!-- badges: end -->

## Overview

The `BiocManager` package, as the modern successor package to `BiocInstaller`,
allows users to install and manage packages from the _[Bioconductor][1]_
project. Bioconductor focuses on the statistical analysis and comprehension of
high-throughput genomic data.

Current _Bioconductor_ packages are available on a 'release' version intended
for every-day use, and a 'devel' version where new features are continually
introduced. A new release version is created every six months. Using the
`BiocManager` package helps users accurately install packages from the
appropriate release.

* `available()` shows all packages associated with a search pattern
* `install()` installs and/or updates packages either CRAN or Bioconductor
* `repositories()` shows all package repository URL endpoints
* `valid()` checks and returns packages that are out-of-date or too new
* `version()` returns the current Bioconductor version number

## Installation

```{r, eval = FALSE}
install.packages("BiocManager")
```

## Usage

```{r, message=FALSE}
BiocManager::version()

BiocManager::install(c("GenomicRanges", "SummarizedExperiment"))

BiocManager::valid()
```

## More information

Please see the package vignette for more detailed information such as changing
Bioconductor version, offline use, and other advanced usage.

## Getting help

To report apparent bugs, create a minimal and reproducible example on
[github][2].

[1]: https://bioconductor.org
[2]: https://github.com/Bioconductor/BiocManager/issues
