---
title: "Installing and Managing _Bioconductor_ Packages"
author:
- name: Marcel Ramos
  affiliation: Roswell Park Cancer Institute, Buffalo, NY
- name: Martin Morgan
  affiliation: Roswell Park Cancer Institute, Buffalo, NY
output:
  BiocStyle::html_document:
      toc: true
vignette: |
  %\VignetteIndexEntry{Installing and Managing Bioconductor Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

Use the [Bioconductor][1] package to install and manage packages from the _[Bioconductor][2]_ project for the statistical analysis and comprehension of high-throughput genomic data.

Current _Bioconductor_ packages are available on a 'release' version intended for every-day use, and a 'devel' version where new features are introduced. A new release version is created every six months. Using the [Bioconductor][1] package helps users install packages from the same release.

# Basic use

Use the standard _R_ installation procedure To install the [Bioconductor][1] package

```{r, eval=FALSE}
chooseCRANmirror()
install.packages("Bioconductor")
```

Then install _Bioconductor_ (or CRAN) packages

```{r, eval=FALSE}
Bioconductor::install(c("GenomicRanges", "Organism.dplyr"))
```

Use `version()` to discover the version of _Bioconductor_ currently in
use.

_Bioconductor_ packages work best when they are all from the same release. Use `valid()` to identify packages that are out-of-date or from unexpected versions.

Questions about installing and managing _Bioconductor_ packages should be addressed to the [_Bioconductor_ support site][4].

# Advanced use

## Changing version

Use the `version=` argument to update all packages to a specific _Bioconductor_ version

```{r, eval=FALSE}
Bioconductor::install(version="3.7")
```

_Bioconductor_ versions are associated with specific _R_ versions, as summarized [here][5]. To use the most recent version of _Bioconductor_ may require installing a new version of _R_.

## Managing multiple versions

It is possible to have multiple versions of _Bioconductor_ installed on the same computer. A best practice is to [create an initial _R_ installation][6]. Then create and use a library for each version of _Bioconductor_. The library will contain all _Bioconductor_, CRAN, and other packages for that version of _Bioconductor_. We illustrate the process assuming use of _Bioconductor_ version 3.7, available using _R_ version 3.5

Create a directory to contain the library (replace `USER_NAME` with your user name on Windows)

- Linux: `~/R/3.5-Bioc-3.7`
- macOS: `~/Library/R/3.5-Bioc-3.7/library`
- Windows: `C:\Users\USER_NAME\Documents\R\3.5-Bioc-3.7`

Set the environment variable `R_LIBS_USER` to this directory, and invoke _R_. Command line examples for Linux are

- Linux: `R_LIBS_USER=~/R/3.5-Bioc-3.7 R`
- macOS: `R_LIBS_USER=~~/Library/R/3.5-Bioc-3.7/library R`
- Windows: `cmd /C "set R_LIBS_USER=C:\Users\USER_NAME\Documents\R\3.5-Bioc-3.7 && R"`

Once in _R_, confirm that the version-specific library path has been set

```{r, eval=FALSE}
.libPaths()
```

On Linux and macOS, create a bash alias to save typing, e.g.,

- Linux: `alias Bioc3.7='R_LIBS_USER=~/R/3.5-Bioc-3.7 R'`
- macOS: `alias Bioc3.7='R_LIBS_USER=~/Library/R/3.5-Bioc-3.7/library R'`

Invoke these from the command line as `Bioc3.7`.

On Windows, create a shortcut. Go to My Computer and navigate to a directory that is in your PATH. Then right-click and choose New->Shortcut.
In the "type the location of the item" box, put:

```
cmd /C "set R_LIBS_USER=C:\Users\USER_NAME\Documents\R\3.5-Bioc-3.7 && R"
```

Click "Next". In the "Type a name for this shortcut" box, type `Bioc-3.7`.

[1]: https://bioconductor.org/packages/Bioconductor
[2]: https://bioconductor.org
[3]: https://support.bioconductor.org
[4]: https://github.com/Bioconductor/Bioconductor
[5]: https://bioconductor.org/about/release-announcements/
[6]: https://cloud.r-project.org/
